### 2021-02-01 README.Rmd
Fix codecov badge

---
title: "camprotR"
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/CambridgeCentreForProteomics/camprotR/workflows/R-CMD-check/badge.svg)](https://github.com/CambridgeCentreForProteomics/camprotR/actions)
[![codecov](https://codecov.io/gh/CambridgeCentreForProteomics/camprotR/branch/master/graph/badge.svg)](https://codecov.io/gh/CambridgeCentreForProteomics/camprotR?branch=master)
<!-- badges: end -->